---
title: "Project 2"
author: "Group 5"
date: "11/10/2019"
output: html_document
---

# Version One

# library packages
```{r}
library(dplyr)
library(dataQualityR)
library(readr)
library(tidyr)
library(randomForest)
library(DMwR)
library(ROSE)
library(pROC)
```

# Data Cleaning
# load data
```{r}
s_asmt <- read.csv("studentAssessment.csv")
asmt <- read_csv("assessments.csv") 
studentVle <- read_csv("studentVle.csv")
info <- read_csv("studentInfo.csv")
```

# clean studentVle (sum_click)
```{r}
# clean the outlier 
median <- median(studentVle$sum_click)
mean <- mean(studentVle$sum_click)
sd <- sd(studentVle$sum_click)
studentVle <- subset(studentVle, studentVle$sum_click < (3 * sd + mean) & studentVle$sum_click > (-3 * sd + mean))

# average the sum_click for each student in each module and presentation
student_vle_group <- studentVle %>%
  group_by(id_student, code_module, code_presentation) %>%
  summarize(sum_click = mean(sum_click))
```

# student_info 
```{r}
info <- info[, -1]
info$imd_band[info$imd_band =="20-Oct"] <- as.character("10-20%")
info$imd_band <- as.factor(info$imd_band)
info$age_band <- as.factor(info$age_band)
info$disability <- as.factor(info$disability)
info$code_module <- as.factor(info$code_module)
info$code_presentation <- as.factor(info$code_presentation)
info$gender <- as.factor(info$gender)
info$region <- as.factor(info$region)
info$highest_education <- as.factor(info$highest_education)
info$num_of_prev_attempts <- as.factor(info$num_of_prev_attempts)
info$studied_credits <- as.factor(info$studied_credits)
info$final_result <- as.factor(info$final_result)
info <- subset(info, final_result != "Withdrawn") # delete withdrawn student
info$final_result <- factor(info$final_result) # drop "Withdrawn" level
info$final_result[info$final_result == "Distinction"] <- "Pass"
info$final_result <- factor(info$final_result) # drop "Distinction" level
```

# join data (first)
```{r}
# first join std_asmt and asmt to get weight and score
student_assessment <- left_join(s_asmt, asmt, by.x = "id_assessment", by.y = "id_assessment")
student_assessment$code_module <- as.factor(student_assessment$code_module)
student_assessment$code_presentation <- as.factor(student_assessment$code_presentation)
# then get final result 
pf_asmt <- left_join(info, student_assessment, by.x = "id_student", by.y = "id_student")
pf_asmt = na.omit(pf_asmt)  
```

# calculate avg score and weight 
```{r}
# for each student, module, and each presentation, find the avg score and weight of each asmt_type
pf_asmt1 <- pf_asmt %>%
  group_by(id_student, code_module, code_presentation, assessment_type) %>%
  summarize(score = mean(score), weight = mean(weight))
```

# join data (second)
```{r}
pf_asmt1$code_module <- as.factor(pf_asmt1$code_module)
pf_asmt1$code_presentation <- as.factor(pf_asmt1$code_presentation)
student_vle_group$code_module <- as.factor(student_vle_group$code_module)
student_vle_group$code_presentation <- as.factor(student_vle_group$code_presentation)

# add score and weight into info table
df1 <- left_join(info, pf_asmt1, by = c("id_student", "code_module", "code_presentation"))
# add avg click 
df2 <- left_join(df1, student_vle_group, by = c("id_student","code_module","code_presentation"))
#df3 <- na.omit(df2)
df3 <- drop_na(df2)
```

# Modeling
# preparing for model
```{r}
# select variables in the df3
dataset <- select(df3, 
                  code_module, code_presentation, gender, region, highest_education, imd_band,
                  studied_credits, num_of_prev_attempts, age_band, disability, assessment_type,
                  sum_click, weight, score, final_result)

set.seed(12345) 
split <- (.7)
index <- createDataPartition(dataset$final_result, p = split, list = FALSE)
train.df <- dataset[index, ] # training data
test.df <- dataset[-index, ] # testing data

train.df$studied_credits <- as.integer(train.df$studied_credits)
test.df$studied_credits <- as.integer(test.df$studied_credits)
train.df <- train.df %>% mutate_if(is.character, as.factor)
test.df <- test.df %>% mutate_if(is.character, as.factor)

prop.table(table(train.df$final_result))

train.under <- ovun.sample(final_result~., data = train.df, method = "under", N = 9502)$data
prop.table(table(train.under$final_result))

train.over <- ovun.sample(final_result~., data = train.df, method = "over", N = 28624)$data
prop.table(table(train.under$final_result))

train.both <- ovun.sample(final_result~., data = train.df, method = "both", N = 28624)$data
prop.table(table(train.both$final_result))
```

# random forest
```{r}
rf_model1 <- randomForest(final_result ~ .,
                          data = train.df, 
                          ntree = 100)
rf_predict1 <- predict(rf_model1, test.df)

rf_model2 <- randomForest(final_result ~ .,
                          data = train.under, 
                          ntree = 100)
rf_predict2 <- predict(rf_model2, test.df)

rf_model3 <- randomForest(final_result ~ .,
                          data = train.over, 
                          ntree = 100)
rf_predict3 <- predict(rf_model3, test.df)

rf_model4 <- randomForest(final_result ~ .,
                          data = train.both, 
                          ntree = 100)
rf_predict4 <- predict(rf_model4, test.df)

confusionMatrix(rf_predict1, test.df$final_result, positive = "Pass")
confusionMatrix(rf_predict2, test.df$final_result, positive = "Pass")
confusionMatrix(rf_predict3, test.df$final_result, positive = "Pass")
confusionMatrix(rf_predict4, test.df$final_result, positive = "Pass")
```

```{r}
varImpPlot(rf_model3)

outcomeName <- 'final_result'
predictorNames <- names(dataset)[names(dataset) != outcomeName]

rf.probs <- predict(rf_model3, test.df[, predictorNames], type = "prob")    
rf.plot <- plot(roc(test.df$final_result, rf.probs[, 2]))
auc(test.df$final_result,rf.probs[, 2])
```

# gbm model
```{r}
fitControl <- trainControl(method = "cv", number = 20, sampling = "up", classProbs = TRUE)

#lm <- (train(train.df[,dependent],train.df[,target], method='glm'))
gbm <- train(train.df[,dependent],train.df[,target], method='gbm', trControl = fitControl)
summary(gbm)
  
gbm.predict <- predict(gbm,test.df[,dependent],type="raw")
#lm.predict <- predict(lm,test.df[,dependent],type="raw")
#summary(gbm.predict)

r <- data.frame(Actual = test.df$final_result , Prediction = gbm.predict)
r <- table(r)

gbm.probs <- predict(gbm,test.df[,dependent],type="prob")    
gbm.plot<-plot(roc(test.df$final_result,gbm.probs[,2]))
confusionMatrix(gbm.predict,test.df[,target], positive = "Pass")
multiclass.roc(test.df$final_result, predict(gbm, test.df[,dependent], type= "prob", percent=FALSE))
```

# svm model
```{r}
svm = svm(formula = final_result ~ .,
           data = train.df,
           type = 'C-classification',
           kernel = 'radial')

svm.predict <- predict(svm,test.df[,dependent], probability = TRUE)
confusionMatrix(svm.predict,test.df[,target])
r <- data.frame(Actual = test.df$final_result , Prediction = svm.predict)
r <- table(r)

multiclass.roc(test.df$final_result, predict(svm, test.df[,dependent], type= "prob", percent=FALSE))
roc_obj <- roc(test.df$final_result, predict(dt, test.df[,dependent], type= "vector", percent=FALSE))
auc(roc_obj)
svm.probs <- predict(svm,test.df[,dependent],type="prob") 
```

# dt model
```{r}
dt <-rpart(final_result ~.,data = train.df, method = "class")
dt

dt.predict <-predict(dt,test.df,type="class")
test.df$final_result<-as.factor(test.df1$final_result)
require(caret)
confusionMatrix(dt.predict,test.df$final_result)
r <- data.frame(Actual = test.df$final_result , Prediction = dt.predict)
r <- table(r)
r
precision <- (r[2,2]/(r[2,2] + r[1,2]))
precision

multiclass.roc(test.df$final_result, predict(dt, test.df[,dependent], type= "prob", percent=FALSE))
```

# bayes model
```{r}
bayes <-naiveBayes(final_result ~.,data = train.df, control=rpart.control(minsplit=2, minbucket=1, cp=0.001))

bayes.predict <- predict(bayes,test.df)
test.df$final_result <- as.factor(test.df$final_result)
bayes.predict <-as.factor(bayes.predict)

require(caret)
confusionMatrix(bayes.predict,test.df$final_result)
r <- data.frame(Actual = test.df$final_result , Prediction = bayes.predict)
r <- table(r)

precision <- (r[2,2]/(r[2,2] + r[1,2]))
               
multiclass.roc(test.df$final_result, predict(bayes, test.df[,dependent], type= "raw", percent=FALSE))
```


# predict test data using random forest
```{r}
avg <- pf_asmt %>%
  group_by(code_module, code_presentation) %>%
  summarize(avg_score = mean(score), avg_weight = mean(weight))

test <- read.csv("test.csv")

test1 <- left_join(test, avg, by = c("code_module", "code_presentation"))

test1$score[is.na(test1$score)] <- 200
for (i in 1:nrow(test1)) {
  if (test1$score[i] == 200)
    test1$score[i] = test1$avg_score[i]
  else
    test1$score[i] = test1$score[i] 
}

test1$weight[is.na(test1$weight)] <- 200
for (i in 1:nrow(test1)) {
  if (test1$weight[i] == 200)
    test1$weight[i] = test1$avg_weight[i]
  else
    test1$weight[i] = test1$weight[i] 
}

# check which assessment_type is most likely to be for each module and presentation
type <- pf_asmt %>%
  group_by(code_module, code_presentation, assessment_type) %>%
  summarize(total = n())

type2 <- data.frame("code_module" = c("AAA", "BBB", "CCC", "DDD", "EEE", "FFF", "GGG"), 
                    "assessment_type2" = c("TMA", "TMA", "TMA", "TMA", "TMA", "CMA", "CMA"))

test2 <- left_join(test1, type2, by.x = "code_module", by.y = "code_module")

test2$assessment_type <- as.character(test2$assessment_type)
test2$assessment_type2 <- as.character(test2$assessment_type2)
test2$assessment_type[is.na(test2$assessment_type)] <- "DN"

for (i in 1:nrow(test2)) {
  if (test2$assessment_type[i] == "DN")
    test2$assessment_type[i] = test2$assessment_type2[i]
  else
    test2$assessment_type[i] = test2$assessment_type[i]
}

test2$assessment_type <- as.factor(test2$assessment_type)

test_final <- select(test2, 
                  code_module, code_presentation, gender, region, highest_education,
                  imd_band, studied_credits, num_of_prev_attempts, age_band, disability,
                  assessment_type, sum_click, weight, score)

test_final$imd_band <- as.character(test_final$imd_band)
test_final$imd_band[test_final$imd_band == "20-Oct"] <- "10-20%"
test_final$imd_band[test_final$imd_band == ""] <- "10-20%"
test_final$imd_band <- as.factor(test_final$imd_band)

test_final$num_of_prev_attempts <- as.factor(test_final$num_of_prev_attempts)
levels(test_final$num_of_prev_attempts) <- levels(train.over$num_of_prev_attempts)

test_final$sum_click[is.na(test_final$sum_click)] <- 3

levels(test_final$code_module) <- levels(test.df$code_module)
levels(test_final$code_presentation) <- levels(test.df$code_presentation)
levels(test_final$gender) <- levels(test.df$gender)
levels(test_final$highest_education) <- levels(test.df$highest_education)
levels(test_final$imd_band) <- levels(test.df$imd_band)
levels(test_final$age_band) <- levels(test.df$age_band)
levels(test_final$disability) <- levels(test.df$disability)
levels(test_final$assessment_type) <- levels(test.df$assessment_type)

test_final$sum_click <- as.numeric(test_final$sum_click)

write.csv(test_final,"/Users/Cecilia/Desktop/BA (Tu)/Project 2/test_final.csv")
```

```{r}
test_predict <- predict(rf_model3, test_final)
test_predict <- as.data.frame(test_predict)
write.csv(test_predict,"/Users/Cecilia/Desktop/BA (Tu)/Project 2/test_predict_answer.csv")
```


# Version Two

# load data
```{r}
assessments <- read_csv("assessments.csv")
studentAssessment <- read_csv("studentAssessment.csv")
studentRegistration <- read_csv("studentRegistration.csv")
studentVle <- read_csv("studentVle.csv")
vle <- read_csv("vle.csv")
studentInfo <- read_csv("studentInfo.csv")
studentInfo <-  subset(studentInfo, final_result != "Withdrawn")
info <-  as.data.frame(studentInfo)
```

```{r}
info <- info[, -1]
info$imd_band[info$imd_band =="20-Oct"] <- as.character("10-20%")
info$imd_band <- as.factor(info$imd_band)
info$age_band <- as.factor(info$age_band)
info$disability <- as.factor(info$disability)
info$code_module <- as.factor(info$code_module)
info$code_presentation <- as.factor(info$code_presentation)
info$gender <- as.factor(info$gender)
info$region <- as.factor(info$region)
info$highest_education <- as.factor(info$highest_education)
info$num_of_prev_attempts <- as.factor(info$num_of_prev_attempts)
info$studied_credits <- as.factor(info$studied_credits)
info$final_result <- as.factor(info$final_result)
info <- subset(info, final_result != "Withdrawn") # delete withdrawn student
info$final_result <- factor(info$final_result) # drop "Withdrawn" level
info$final_result[info$final_result == "Distinction"] <- "Pass"
info$final_result <- factor(info$final_result) # drop "Distinction" level
```

# join data (first)
```{r}
dataset <- left_join(studentAssessment, assessments, by.x = "id_assessment", by.y = "id_assessment" )
dataset1 <- left_join(info, dataset, by.x = "id_student", by.y = "id_student" )
dataset2 <- left_join(studentVle, vle, by.x = "id_site", by.y = "id_site" )
dataset3 <- left_join(dataset1, dataset2, by.x = "id_student", by.y = "id_student" )
dataset3$week_from = NULL
dataset3$week_to = NULL
dataset4 = na.omit(dataset3) # Removing Duplicates
prop.table(table(dataset4$final_result))
```


# preparing for model
```{r}
# select variables in the df3
dataset5 <- select(dataset4, 
                  code_module, code_presentation, gender, region, highest_education, imd_band,
                  studied_credits, num_of_prev_attempts, age_band, disability, assessment_type,activity_type,
                  sum_click, weight, score, final_result)

set.seed(12345)
split <- (.5)
index <- createDataPartition(dataset5$final_result, p = split, list = FALSE)
train.df <- dataset5[index, ] # training data
test.df <- dataset5[-index, ] # testing data

train.df$studied_credits <- as.integer(train.df$studied_credits)
test.df$studied_credits <- as.integer(test.df$studied_credits)
train.df <- train.df %>% mutate_if(is.character, as.factor)
test.df <- test.df %>% mutate_if(is.character, as.factor)

v = table(train.df$final_result)

train.under <- ovun.sample(final_result~., data = train.df, method = "under", N = v[1]*2)$data
prop.table(table(train.under$final_result))

train.over <- ovun.sample(final_result~., data = train.df, method = "over", N = v[2]*2)$data
prop.table(table(train.under$final_result))

train.both <- ovun.sample(final_result~., data = train.df, method = "both", N = v[2]*2)$data
prop.table(table(train.both$final_result))

test.df <- rbind(train.df[1, ] , test.df)
test.df <- test.df[-1,]
```

# random forest
```{r}
rf_model1 <- randomForest(final_result ~ .,
                          data = train.df, 
                          ntree = 100)
summary(rf_model1)
varImpPlot(rf_model1)
rf_predict1 <- predict(rf_model1, test.df)

rf_model2 <- randomForest(final_result ~ .,
                          data = train.under, 
                          ntree = 100)
varImpPlot(rf_model2)
rf_predict2 <- predict(rf_model2, test.df)

rf_model3 <- randomForest(final_result ~ .,
                          data = train.over, 
                          ntree = 100)
varImpPlot(rf_model3)
str(test.df)
rf_predict3 <- predict(rf_model3, test.df)

rf_model4 <- randomForest(final_result ~ .,
                          data = train.both, 
                          ntree = 100)
varImpPlot(rf_model4)
rf_predict4 <- predict(rf_model4, test.df)

multiclass.roc(test.df$final_result, predict(rf_model1, test.df[,1:15], type= "prob", percent=FALSE))
confusionMatrix(rf_predict1, test.df$final_result, positive = "Pass") 
confusionMatrix(rf_predict2, test.df$final_result, positive = "Pass")
confusionMatrix(rf_predict3, test.df$final_result, positive = "Pass")
confusionMatrix(rf_predict4, test.df$final_result, positive = "Pass")
```

```{r}
outcomeName <- 'final_result'
predictorNames <- names(dataset5)[names(dataset5) != outcomeName]

rf.probs <- predict(rf_model2, test.df[, predictorNames], type = "prob")    
rf.plot <- plot(roc(test.df$final_result, rf.probs[, 2]))
auc(test.df$final_result,rf.probs[, 2])
```

# predict test data
```{r}
test_final2 <- select(test2, 
                  code_module, code_presentation, gender, region, highest_education,
                  imd_band, studied_credits, num_of_prev_attempts, age_band, disability,
                  assessment_type, activity_type, sum_click, weight, score)

test_final2$imd_band <- as.character(test_final2$imd_band)
test_final2$imd_band[test_final2$imd_band == "20-Oct"] <- "10-20%"
test_final2$imd_band[test_final2$imd_band == ""] <- "10-20%"
test_final2$imd_band <- as.factor(test_final2$imd_band)

test_final2$num_of_prev_attempts <- as.factor(test_final2$num_of_prev_attempts)
levels(test_final2$num_of_prev_attempts) <- levels(train.over$num_of_prev_attempts)

test_final2$sum_click[is.na(test_final2$sum_click)] <- 3

levels(test_final2$code_module) <- levels(test.df$code_module)
levels(test_final2$code_presentation) <- levels(test.df$code_presentation)
levels(test_final2$gender) <- levels(test.df$gender)
levels(test_final2$highest_education) <- levels(test.df$highest_education)
levels(test_final2$imd_band) <- levels(test.df$imd_band)
levels(test_final2$age_band) <- levels(test.df$age_band)
levels(test_final2$disability) <- levels(test.df$disability)
levels(test_final2$assessment_type) <- levels(test.df$assessment_type)

test_final2$sum_click <- as.numeric(test_final2$sum_click)

test_final2$activity_type[is.na(test_final2$activity_type)] <- "homepage"
```

```{r}
prof_test <- select(test_final2, 
                  code_module, code_presentation, gender, region, highest_education, imd_band,
                  studied_credits, num_of_prev_attempts, age_band, disability, assessment_type,activity_type,
                  sum_click, weight, score)

prof_test$studied_credits <- as.integer(prof_test$studied_credits)
prof_test$studied_credits <- as.integer(prof_test$studied_credits)
prof_test$num_of_prev_attempts <- as.factor(prof_test$num_of_prev_attempts)
prof_test <- prof_test %>% mutate_if(is.character, as.factor)
prof_test <- prof_test %>% mutate_if(is.character, as.factor)

str(prof_test)
str(train.df)
levels(prof_test$activity_type ) <- levels(train.df$activity_type)
levels(prof_test$num_of_prev_attempts ) <- levels(train.df$num_of_prev_attempts)
test_predict <- predict(rf_model4, prof_test)
test_predict <- as.data.frame(test_predict)
write.csv(test_predict,"G:/My Drive/1 SEM/MG-GY-9753 Business Analytics/Project_2/test_predict.csv")
prop.table(table(test_predict))
```






